Comparing Facebook's Prophet Forecasting to an ARIMA Model: A Diebold-Mariano Evaluation of the Top40 Index

### 2017-09-15 template/Draft_Paper.Rmd
Final typo edits

The challenge with many forecasting methods such as the Box-Jenkins methodology and machine-learning approaches are that they usually require a sufficient statistical background. Furthermore, these approaches tend to be inflexible with respect to incorporating prior information. Though this can be partially acheived by initializing parameter estimates for methods such as Artificial Neural Networks ,an analyst with no statistical training can have difficulty interpreting and thus modifying parameter estimates[@taylor2017forecasting].  
Prophet is an **automatic forecasting** tool that uses a Bayesian Generalized Additive model to generate time series forecasts. This is in comparison to conventional methods such as the linear stochastic dependence that exists in ARIMA models. Prophet combines a configurable model that includes performance analysis evaluation with the interaction of an analyst in the loop [@taylor2017forecasting]. The model is configurable which allows an analyst to incorporate their knowledge of the behaviour of the series into the model building process through easily interpretable initial parameters that can be modified and interactive feedback when forecasts under-perform [@taylor2017forecasting].

Through this mechanism, a large numer of forecasts can be reliably generated **automatically** with the flexibility of enabling an analyst to modify the model in-the-loop during model specification. The automated forecasting procedure with an analyst-in-the-loop is illustrated in Figure 2.

@@

The growth component of the model is generalized with two useful features that allow the carrying capacity and growth rate of the model to vary with time. The model enables an analyst to manually define when and how the growth rate changes at different change points [@taylor2017forecasting]. This feature is particularly useful for forecasting stock price data since analysts are able to incorporate market movements that lead to the growth rate either decreasing, increasing or remaining constant.  
The seasonality component, $s(t)$ with periodicity $P$ and is given by,
$$s(t) = \sum_{n = -N}^{N}c_n \exp(j\frac{2\pi nt}{P})$$

@@

$$ h(t) = \sum_{j = 1}^{L}\kappa_j \boldsymbol{1}(t \in D_j) \label{eqn5}
$$
Where, $\kappa$ is normally distributed scaling factor and  $D_j$ is a set of past and future dates where holidays or events occur.

Prophet translates model parameter estimation into a curve-fitting exercise by employing Maximum Apriori Posterior (MAP) estimation, a quasi-Bayesian approach, to estimate the optimal parameter values for the forecasting model. MAP estimation finds the maximum posterior estimates for the parameters by using the likelihood function and an apriori distribution to find a posterior distribution which is optimized to return point estimates of the parameters [@taylor2017forecasting]. Thereafter, the package automatically evaluates the forecasts by comparing the Simulated Historical Forecast (SHF) errors of baseline forecasting methods (including ARIMA forecasts) to those of Prophet. SHF errors are forecast errors generated from random points in the history of the time series model. In Prophet, SHF errors are compared visually to allow the analysts to easily adjust the interpretable parameters before automatically estimating the model and evaluating forecasts.

@@

The data set used throughout this paper comprises of the daily closing prices of the equally-weighted FTSE/JSE Top40 price index for the five year period from 31/07/2012 to 30/07/2017. This equates to a total of 1766 trading days excluding weekends and holidays. This period was chosen for its recency and to avoid exposing the results to the noise associated with the high volatility experienced in the stock market during the 2008 financial crisis. The aim is to compare the predictive accuracy of the two models under standard market conditions before other confounding factors can be closely analysed.  Figure 33 shows the actual realisations of the price index which was sourced from Datastream.

@@

The price index was transformation to log returns as a means of avoid spurious regression issues which can be caused by the existence of unit roots in stock prices.

@@

Preliminary data processing was done to clean and transform the data before formally testing for outliers. The data was then split into in-sample and out-of-sample periods. The in-sample period comprises of 1400 trading days from 30 July 2012 to the 1 May 2016. The out_of_sample period comprises of the remaining 366 trading days from the dataset.

@@

To identify what ARIMA model to use, the methodology as outlined in @box1970time was applied to the data. Firstly, stationarity conditions were verified formally by performing an augmented Dickey-Fuller test on the in-sample returns data. The results of the test suggest the p_value is less that 1%, which implies that the in-sample returns series is stationary. The results in the autocorrelation(ACF) and partial autocorrelation(PACF) function illustrated in figure 4  suggest that the second lag of both functions might be significantly different from zero.  
![](acf & pacf plot.png){ width=100%}  
*Figure 4 : ACF and PACF plots for in-sample data*

Therefore, the candidate *ARIMA(p,d,q)* models in Table 1 were fit to the in-sample data and the AIC used to choose the model that best fits the data. In Table 1, the candidate model with the lowest AIC is the $ARIMA(2,0,2)$. This model was thus selected as the ARIMA forecasting model.  
![](AIC.png){ width=100% }  
*Table 1: AIC for Candidate ARIMA models*

## Forecast Error Evaluation
The results of model estimation and forecast evaluation using the rolling window for h-step forecasts yielded forecasts errors with roughly equal distributions. This is visible from the boxplot in Figure 5. The presence of outliers seems consistent across the different models and forecastign horizons. Furthermore, the variation in the accuracy across different forecasting horizons for each model are not observable in Figure 5. The mean error and spread of both models across different time horizons appears to be constant. 
![](Forecast error boxplot.png){ width=105% }  
*Figure 5: Boxplot illustrating distribution of rolling forecast errors for different time horizons, h*

@@

Table 2 summarizes the forecast error statistics for both the ARIMA and Prophet model across the three different time horizons.  
Across all forecast horizons, the ARIMA model appears to be overestimating returns. This is evident from the negative mean error(ME) across $h = (1,5,20)$. On the contrary, the Prophet model underestimates the forecasts for one-step and five-step ahead forecasts. It then overstimates returns when forecasting twenty-steps ahead. The ME results shown in Table 2 fail to clearly illustrate which model performs better. Furthermore, the ME values are relatively small when compared against other error statistics.

@@

The observations made from Table 2 in section 5.2 suggest that the ARIMA and Prophet models have equal predictive accuracy. These assertions can be formally quantified through a Diebold-Mariano(DM) test which evaluates whether the models have equal predictive accuracy against the null hypothesis that one model has superior accuracy over the other. Table 3 reports the results obtained when the rolling forecast errors were formally compared against each other over different time horizons.  
![](dm results.png){ width=105%}  
*Table 3 Results of the Diebold-Mariano Hypothesis Test*  
The P-value across all the time horizons for either linear and quadratic loss functions suggest that we would fail to reject the null hypothesis which implies that the predictive accuracy of the models could be equal and the 95% level. This observation validates the results obtained in section 5.2.

#Discussion of Results
The results of the study show that there is no significant evidence to conclude that Prophet generates superior forecasts for returns on the FTSE/JSE Top40 Index when compared against those of an ARIMA model when the Diebold-Mariano Test is used as a method of evaluation. These results are consistent for daily, weekly and monthly forecasting horizons.  
Nevertheless, the study has flagged the possibility that, on average, the ARIMA model tends to overestimate forecasted returns on the index for the forecast horizons under consideration. This is contrary to Prophet, that, on average, underestimates forecasted returns over daily and weekly time horizons. This information could be useful for an analyst engaged in short-term strategies since the choice of the forecasting model can reflect a particular sentiment or risk appetite. For example, if the analyst feels like prevailing returns are underpriced by the market, they might choose to use an ARIMA model, whereas an opposite sentiment could warrant the use pf Prophet.  
Furthermore, it is worth noting that the merits of Prophet may have not been fully exposed in this study due to the choice of the forecasting horizon, i.e. excluding periods of high volatility and uncertainity during and after the 2008 Financial Crisis. Nonetheless, it is reassuring that the model has performed as well as the ARIMA model. This observation could create room for the application of Prophet in circumstances where flexibility in the model formulation is desireable.

@@

The challenge of building models that generate superior stock return forecasts has led to widespread research in the field of quantitative analysis for decades, but the ARIMA model remains the standard against which to compare new innovations. This is because of its relative simplicity and ability to evolve into more complex applications e.g in the hybrid models considered in the literature review.  
Prophet has reduced the problem of estimating a time series model to a curve fittig exercise by defining an additive, parametric function of time. The functions' parameters are then estimated using Bayesian estimation methods after which the tool automatically executes forecast evaluation to optimize its predictive ability.  
The results of this study suggest that, using a Diebold-Mariano evaluation, ARIMA and Prophet do not have a significant difference in predictive accuracy, even across time horizons of up to a month. The implications of this are that analysts can then chosse which model to use based on their needs without significantly compromising forecast  accuracy.  
This study was confined to the analysis of predictive performance during general market conditions. Further research could explore the predictive ability of the models during different market conditions.For example, performance can be compared during bull and bear market conditions or during periods of high volatility. Furthermore, research can assess whether prophet would perform better on raw data that has not been transformed, since the model is built with growth and seasonality components and could thus predict better forecasts if these features are exploited.

>>>>

The challenge with many forecasting methods such as the Box-Jenkins methodology and machine-learning approaches are that they usually require a sufficient statistical background. Furthermore, these approaches tend to be inflexible with respect to incorporating prior information. Though this can be partially achieved by initializing parameter estimates for some methods (such as ANNs) ,an analyst with no statistical training can have difficulty interpreting and thus modifying parameter estimates[@taylor2017forecasting].  
Prophet is an **automatic forecasting** tool that uses a Bayesian Generalized Additive model to generate time series forecasts. This is in comparison to conventional methods such as the linear stochastic dependence that exists in ARIMA models. Prophet combines a configurable model that includes performance analysis evaluation with the interaction of an analyst in the loop [@taylor2017forecasting]. The model is configurable which allows an analyst to incorporate their knowledge of the behaviour of the series into the model building process through easily interpretable initial parameters that can be modified and interactive feedback when forecasts under-perform [@taylor2017forecasting].

Through this mechanism, a large number of forecasts can be reliably generated **automatically** with the flexibility of enabling an analyst to modify the model in-the-loop during model specification. The automated forecasting procedure with an analyst-in-the-loop is illustrated in Figure 2.

@@

The growth component of the model has two useful features that allow the carrying capacity and growth rate of the model to vary with time. This enables an analyst to manually define when and how the growth rate changes at different change points [@taylor2017forecasting]. This feature is particularly useful for forecasting stock price data since analysts are able to incorporate market movements that lead to the growth rate either decreasing, increasing or becoming constant.  
The seasonality component, $s(t)$ with periodicity $P$ is given by,
$$s(t) = \sum_{n = -N}^{N}c_n \exp(j\frac{2\pi nt}{P})$$

@@

$$ h(t) = \sum_{j = 1}^{L}\kappa_j \boldsymbol{1}(t \in D_j) \label{eqn5}
$$
Where, $\kappa$ is a normally distributed scaling factor and  $D_j$ is a set of past and future dates where holidays or events occur.

Prophet translates model parameter estimation into a curve-fitting exercise by employing Maximum Apriori Posterior (MAP) estimation, a Bayesian approach to estimating optimal parameter values for the forecasting model. MAP estimation finds the maximum posterior estimates for the parameters by using the likelihood function and an apriori distribution to evaluate a posterior distribution which is then optimized to return point estimates of the parameters [@taylor2017forecasting]. Thereafter, the package automatically evaluates the forecasts by comparing the Simulated Historical Forecast (SHF) errors of baseline forecasting methods (including ARIMA forecasts) to those of Prophet. SHF errors are forecast errors generated from random points in the history of the time series model. In Prophet, SHF errors are compared visually to allow the analysts to easily adjust the interpretable parameters before automatically estimating the model and evaluating forecasts.

@@

The data set used throughout this paper comprises of the daily closing prices of the equally-weighted FTSE/JSE Top40 price index for the five year period from 31/07/2012 to 30/07/2017. This equates to a total of 1766 trading days excluding weekends and holidays. This period was chosen for its recency and to avoid exposing the results to the noise associated with the high volatility experienced in the stock market during the 2008 financial crisis. The aim is to compare the predictive accuracy of the two models under standard market conditions before other confounding factors can be closely analysed.  Figure 3 shows the actual realisations of the price index data which was sourced from Datastream.

@@

The price index was transformed to log returns as a means of avoiding spurious regression issues which can be caused by the existence of unit roots in stock price data.

@@

Preliminary data processing was done to clean and transform the data before formally testing for outliers. The data was then split into in-sample and out-of-sample periods. The in-sample period comprises of 1400 trading days from 30 July 2012 to the 1 May 2016. The out-of-sample period comprises of the remaining 366 trading days from the dataset.

@@

To identify what ARIMA model to use, the methodology as outlined in @box1970time was applied to the data. Firstly, stationarity conditions were verified formally by performing an Augmented Dickey-Fuller test on the in-sample returns data. The results of the test suggest the p_value is less that 1%, which implies that the in-sample returns series is stationary. The results in the autocorrelation(ACF) and partial autocorrelation(PACF) function illustrated in Figure 4  suggest that the second lag of both functions might be significantly different from zero.  
![](acf & pacf plot.png){ width=100%}  
*Figure 4 : ACF and PACF plots for in-sample data*

Observations from the above plots informed the choice of candidate *ARIMA(p,d,q)* models to fit to the data. The candidate models that were fit to the in-sample data are shown in Table 1 with their respective AIC values. The candidate model with the lowest AIC is the $ARIMA(2,0,2)$. This model was thus selected as the ARIMA forecasting model.  
![](AIC.png){ width=100% }  
*Table 1: AIC for Candidate ARIMA models*

## Forecast Error Evaluation
The results of model estimation and forecast evaluation using the rolling window for h-step forecasts yielded forecasts errors with roughly equal distributions. This is visible from the boxplot in Figure 5. The presence of outliers seems consistent across the different models **and** forecasting horizons. Furthermore, variations in accuracy across different forecasting horizons *for each* model is not observable in Figure 5. The mean error and spread of both models across different time horizons appears to be constant. 
![](Forecast error boxplot.png){ width=105% }  
*Figure 5: Boxplot illustrating distribution of rolling forecast errors for different time horizons, h*

@@

Table 2 summarizes the forecast error statistics for both the ARIMA and Prophet model across the three different time horizons.  
Across all forecast horizons, the ARIMA model appears to be overestimating returns. This is evident from the negative mean error(ME) across $h = (1,5,20)$. On the contrary, the Prophet model underestimates the forecasts for one-step and five-step ahead forecasting horizons. It then overstimates returns when forecasting twenty-steps ahead. The ME results shown in Table 2 fail to clearly illustrate which model performs better. Furthermore, the ME values are relatively small when compared against other error statistics.

@@

The observations made from Table 2 in section 5.2.1 suggest that the ARIMA and Prophet models have equal predictive accuracy. These assertions can be formally quantified through a Diebold-Mariano(DM) test which evaluates whether the models have equal predictive accuracy against the alternative hypothesis that one model has superior accuracy over the other. Table 3 reports the results obtained when the rolling forecast errors were formally compared against each other over different time horizons.  
![](dm results.png){ width=105%}  
*Table 3 Results of the Diebold-Mariano Hypothesis Test*  
The p-value across all the time horizons for either linear and quadratic loss functions suggest that we would fail to reject the null hypothesis. This implies that the predictive accuracy of the models could be equal. These results formally validate the observations made obtained in section 5.2.1.

#Discussion of Results
The results of the study show that there is no significant evidence to conclude that Prophet generates superior forecasts for returns on the FTSE/JSE Top40 Index when compared against those of an ARIMA model when the Diebold-Mariano Test is used as a method of evaluation. These results are consistent for daily, weekly and monthly forecasting horizons.  
Nevertheless, the study has flagged the possibility that, on average, the ARIMA model tends to overestimate forecasted returns on the index for the forecast horizons under consideration. This is contrary to Prophet, that, on average, underestimates forecasted returns over daily and weekly time horizons. This information could be useful for an analyst engaged in short-term strategies since the choice of the forecasting model can reflect a particular sentiment or risk appetite. For example, if the analyst feels like prevailing returns are underpriced by the market, they might choose to use an ARIMA model, whereas an opposite sentiment could warrant the use of Prophet.  
Furthermore, it is worth noting that the merits of Prophet may have not been fully exposed in this study due to the use of transformed data and the choice of a forecasting horizon that excluded periods of high volatility. Nonetheless, it is reassuring that the model has performed as well as the ARIMA model. This observation could create room for the application of Prophet in circumstances where flexibility in the model formulation is desireable.

@@

The challenge of building models that generate superior stock return forecasts has led to widespread research in the field of quantitative analysis for decades, but the ARIMA model remains the standard against which to compare new innovations. This is because of its relative simplicity and ability to evolve into more complex applications e.g in the hybrid models considered in the literature review.  
Prophet has reduced the problem of estimating a time series model to a curve fittig exercise by defining an additive, parametric function of time. The functions' parameters are then estimated using Bayesian estimation methods after which the tool automatically executes forecast evaluation to optimize its predictive ability.  
The results of this study suggest that, using a Diebold-Mariano evaluation, the ARIMA and Prophet forecasting models do not have a significant difference in predictive accuracy across time horizons of up to a month. The implications of this are that analysts can then choose which model to use based on their needs without significantly compromising forecast  accuracy.  
This study was confined to the analysis of predictive performance during general market conditions. Further research could explore the predictive ability of the models during different market conditions. For example, performance can be compared during bull and bear market conditions or during periods of high volatility. Additionally, further research can assess whether prophet would perform better on raw data that has not been transformed, since the model is built with growth and seasonality components. This use of which could lead to Prophet models that generate superior forecasts when compared against alternatives.

### 2017-09-15 template/Draft_Paper.tex
Final typo edits

The challenge with many forecasting methods such as the Box-Jenkins
methodology and machine-learning approaches are that they usually
require a sufficient statistical background. Furthermore, these
approaches tend to be inflexible with respect to incorporating prior
information. Though this can be partially acheived by initializing
parameter estimates for methods such as Artificial Neural Networks ,an
analyst with no statistical training can have difficulty interpreting
and thus modifying parameter estimates(Taylor and Letham 2017).\\
Prophet is an \textbf{automatic forecasting} tool that uses a Bayesian
Generalized Additive model to generate time series forecasts. This is in
comparison to conventional methods such as the linear stochastic
dependence that exists in ARIMA models. Prophet combines a configurable
model that includes performance analysis evaluation with the interaction
of an analyst in the loop (Taylor and Letham 2017). The model is
configurable which allows an analyst to incorporate their knowledge of
the behaviour of the series into the model building process through
easily interpretable initial parameters that can be modified and
interactive feedback when forecasts under-perform (Taylor and Letham
2017).

Through this mechanism, a large numer of forecasts can be reliably
generated \textbf{automatically} with the flexibility of enabling an
analyst to modify the model in-the-loop during model specification. The
automated forecasting procedure with an analyst-in-the-loop is
illustrated in Figure 2.

@@

The growth component of the model is generalized with two useful
features that allow the carrying capacity and growth rate of the model
to vary with time. The model enables an analyst to manually define when
and how the growth rate changes at different change points (Taylor and
Letham 2017). This feature is particularly useful for forecasting stock
price data since analysts are able to incorporate market movements that
lead to the growth rate either decreasing, increasing or remaining
constant.\\
The seasonality component, \(s(t)\) with periodicity \(P\) and is given
by, \[s(t) = \sum_{n = -N}^{N}c_n \exp(j\frac{2\pi nt}{P})\]

@@

\[ h(t) = \sum_{j = 1}^{L}\kappa_j \boldsymbol{1}(t \in D_j) \label{eqn5}
\] Where, \(\kappa\) is normally distributed scaling factor and \(D_j\)
is a set of past and future dates where holidays or events occur.

Prophet translates model parameter estimation into a curve-fitting
exercise by employing Maximum Apriori Posterior (MAP) estimation, a
quasi-Bayesian approach, to estimate the optimal parameter values for
the forecasting model. MAP estimation finds the maximum posterior
estimates for the parameters by using the likelihood function and an
apriori distribution to find a posterior distribution which is optimized
to return point estimates of the parameters (Taylor and Letham 2017).
Thereafter, the package automatically evaluates the forecasts by
comparing the Simulated Historical Forecast (SHF) errors of baseline
forecasting methods (including ARIMA forecasts) to those of Prophet. SHF
errors are forecast errors generated from random points in the history
of the time series model. In Prophet, SHF errors are compared visually
to allow the analysts to easily adjust the interpretable parameters
before automatically estimating the model and evaluating forecasts.

@@

Preliminary data processing was done to clean and transform the data
before formally testing for outliers. The data was then split into
in-sample and out-of-sample periods. The in-sample period comprises of
1400 trading days from 30 July 2012 to the 1 May 2016. The
out\_of\_sample period comprises of the remaining 366 trading days from
the dataset.

@@

To identify what ARIMA model to use, the methodology as outlined in Box
and Jenkins (1970) was applied to the data. Firstly, stationarity
conditions were verified formally by performing an augmented
Dickey-Fuller test on the in-sample returns data. The results of the
test suggest the p\_value is less that 1\%, which implies that the
in-sample returns series is stationary. The results in the
autocorrelation(ACF) and partial autocorrelation(PACF) function
illustrated in figure 4 suggest that the second lag of both functions
might be significantly different from zero.\\
\includegraphics[width=1.00000\textwidth]{acf \& pacf plot.png}\\
\emph{Figure 4 : ACF and PACF plots for in-sample data}

Therefore, the candidate \emph{ARIMA(p,d,q)} models in Table 1 were fit
to the in-sample data and the AIC used to choose the model that best
fits the data. In Table 1, the candidate model with the lowest AIC is
the \(ARIMA(2,0,2)\). This model was thus selected as the ARIMA
forecasting model.\\
\includegraphics[width=1.00000\textwidth]{AIC.png}\\
\emph{Table 1: AIC for Candidate ARIMA models}

@@

The results of model estimation and forecast evaluation using the
rolling window for h-step forecasts yielded forecasts errors with
roughly equal distributions. This is visible from the boxplot in Figure
5. The presence of outliers seems consistent across the different models
and forecastign horizons. Furthermore, the variation in the accuracy
across different forecasting horizons for each model are not observable
in Figure 5. The mean error and spread of both models across different
time horizons appears to be constant.
\includegraphics[width=1.05000\textwidth]{Forecast error boxplot.png}\\
\emph{Figure 5: Boxplot illustrating distribution of rolling forecast
errors for different time horizons, h}

@@

Table 2 summarizes the forecast error statistics for both the ARIMA and
Prophet model across the three different time horizons.\\
Across all forecast horizons, the ARIMA model appears to be
overestimating returns. This is evident from the negative mean error(ME)
across \(h = (1,5,20)\). On the contrary, the Prophet model
underestimates the forecasts for one-step and five-step ahead forecasts.
It then overstimates returns when forecasting twenty-steps ahead. The ME
results shown in Table 2 fail to clearly illustrate which model performs
better. Furthermore, the ME values are relatively small when compared
against other error statistics.

@@

The observations made from Table 2 in section 5.2 suggest that the ARIMA
and Prophet models have equal predictive accuracy. These assertions can
be formally quantified through a Diebold-Mariano(DM) test which
evaluates whether the models have equal predictive accuracy against the
null hypothesis that one model has superior accuracy over the other.
Table 3 reports the results obtained when the rolling forecast errors
were formally compared against each other over different time
horizons.\\
\includegraphics[width=1.05000\textwidth]{dm results.png}\\
\emph{Table 3 Results of the Diebold-Mariano Hypothesis Test}\\
The P-value across all the time horizons for either linear and quadratic
loss functions suggest that we would fail to reject the null hypothesis
which implies that the predictive accuracy of the models could be equal
and the 95\% level. This observation validates the results obtained in
section 5.2.

@@

The results of the study show that there is no significant evidence to
conclude that Prophet generates superior forecasts for returns on the
FTSE/JSE Top40 Index when compared against those of an ARIMA model when
the Diebold-Mariano Test is used as a method of evaluation. These
results are consistent for daily, weekly and monthly forecasting
horizons.\\
Nevertheless, the study has flagged the possibility that, on average,
the ARIMA model tends to overestimate forecasted returns on the index
for the forecast horizons under consideration. This is contrary to
Prophet, that, on average, underestimates forecasted returns over daily
and weekly time horizons. This information could be useful for an
analyst engaged in short-term strategies since the choice of the
forecasting model can reflect a particular sentiment or risk appetite.
For example, if the analyst feels like prevailing returns are
underpriced by the market, they might choose to use an ARIMA model,
whereas an opposite sentiment could warrant the use pf Prophet.\\
Furthermore, it is worth noting that the merits of Prophet may have not
been fully exposed in this study due to the choice of the forecasting
horizon, i.e.~excluding periods of high volatility and uncertainity
during and after the 2008 Financial Crisis. Nonetheless, it is
reassuring that the model has performed as well as the ARIMA model. This
observation could create room for the application of Prophet in
circumstances where flexibility in the model formulation is desireable.

@@

The challenge of building models that generate superior stock return
forecasts has led to widespread research in the field of quantitative
analysis for decades, but the ARIMA model remains the standard against
which to compare new innovations. This is because of its relative
simplicity and ability to evolve into more complex applications e.g in
the hybrid models considered in the literature review.\\
Prophet has reduced the problem of estimating a time series model to a
curve fittig exercise by defining an additive, parametric function of
time. The functions' parameters are then estimated using Bayesian
estimation methods after which the tool automatically executes forecast
evaluation to optimize its predictive ability.\\
The results of this study suggest that, using a Diebold-Mariano
evaluation, ARIMA and Prophet do not have a significant difference in
predictive accuracy, even across time horizons of up to a month. The
implications of this are that analysts can then chosse which model to
use based on their needs without significantly compromising forecast
accuracy.\\
This study was confined to the analysis of predictive performance during
general market conditions. Further research could explore the predictive
ability of the models during different market conditions.For example,
performance can be compared during bull and bear market conditions or
during periods of high volatility. Furthermore, research can assess
whether prophet would perform better on raw data that has not been
transformed, since the model is built with growth and seasonality
components and could thus predict better forecasts if these features are
exploited.

>>>>

The challenge with many forecasting methods such as the Box-Jenkins
methodology and machine-learning approaches are that they usually
require a sufficient statistical background. Furthermore, these
approaches tend to be inflexible with respect to incorporating prior
information. Though this can be partially achieved by initializing
parameter estimates for some methods (such as ANNs) ,an analyst with no
statistical training can have difficulty interpreting and thus modifying
parameter estimates(Taylor and Letham 2017).\\
Prophet is an \textbf{automatic forecasting} tool that uses a Bayesian
Generalized Additive model to generate time series forecasts. This is in
comparison to conventional methods such as the linear stochastic
dependence that exists in ARIMA models. Prophet combines a configurable
model that includes performance analysis evaluation with the interaction
of an analyst in the loop (Taylor and Letham 2017). The model is
configurable which allows an analyst to incorporate their knowledge of
the behaviour of the series into the model building process through
easily interpretable initial parameters that can be modified and
interactive feedback when forecasts under-perform (Taylor and Letham
2017).

Through this mechanism, a large number of forecasts can be reliably
generated \textbf{automatically} with the flexibility of enabling an
analyst to modify the model in-the-loop during model specification. The
automated forecasting procedure with an analyst-in-the-loop is
illustrated in Figure 2.

@@

The growth component of the model has two useful features that allow the
carrying capacity and growth rate of the model to vary with time. This
enables an analyst to manually define when and how the growth rate
changes at different change points (Taylor and Letham 2017). This
feature is particularly useful for forecasting stock price data since
analysts are able to incorporate market movements that lead to the
growth rate either decreasing, increasing or becoming constant.\\
The seasonality component, \(s(t)\) with periodicity \(P\) is given by,
\[s(t) = \sum_{n = -N}^{N}c_n \exp(j\frac{2\pi nt}{P})\]

@@

\[ h(t) = \sum_{j = 1}^{L}\kappa_j \boldsymbol{1}(t \in D_j) \label{eqn5}
\] Where, \(\kappa\) is a normally distributed scaling factor and
\(D_j\) is a set of past and future dates where holidays or events
occur.

Prophet translates model parameter estimation into a curve-fitting
exercise by employing Maximum Apriori Posterior (MAP) estimation, a
Bayesian approach to estimating optimal parameter values for the
forecasting model. MAP estimation finds the maximum posterior estimates
for the parameters by using the likelihood function and an apriori
distribution to evaluate a posterior distribution which is then
optimized to return point estimates of the parameters (Taylor and Letham
2017). Thereafter, the package automatically evaluates the forecasts by
comparing the Simulated Historical Forecast (SHF) errors of baseline
forecasting methods (including ARIMA forecasts) to those of Prophet. SHF
errors are forecast errors generated from random points in the history
of the time series model. In Prophet, SHF errors are compared visually
to allow the analysts to easily adjust the interpretable parameters
before automatically estimating the model and evaluating forecasts.

@@

Preliminary data processing was done to clean and transform the data
before formally testing for outliers. The data was then split into
in-sample and out-of-sample periods. The in-sample period comprises of
1400 trading days from 30 July 2012 to the 1 May 2016. The out-of-sample
period comprises of the remaining 366 trading days from the dataset.

@@

To identify what ARIMA model to use, the methodology as outlined in Box
and Jenkins (1970) was applied to the data. Firstly, stationarity
conditions were verified formally by performing an Augmented
Dickey-Fuller test on the in-sample returns data. The results of the
test suggest the p\_value is less that 1\%, which implies that the
in-sample returns series is stationary. The results in the
autocorrelation(ACF) and partial autocorrelation(PACF) function
illustrated in Figure 4 suggest that the second lag of both functions
might be significantly different from zero.\\
\includegraphics[width=1.00000\textwidth]{acf \& pacf plot.png}\\
\emph{Figure 4 : ACF and PACF plots for in-sample data}

Observations from the above plots informed the choice of candidate
\emph{ARIMA(p,d,q)} models to fit to the data. The candidate models that
were fit to the in-sample data are shown in Table 1 with their
respective AIC values. The candidate model with the lowest AIC is the
\(ARIMA(2,0,2)\). This model was thus selected as the ARIMA forecasting
model.\\
\includegraphics[width=1.00000\textwidth]{AIC.png}\\
\emph{Table 1: AIC for Candidate ARIMA models}

@@

The results of model estimation and forecast evaluation using the
rolling window for h-step forecasts yielded forecasts errors with
roughly equal distributions. This is visible from the boxplot in Figure
5. The presence of outliers seems consistent across the different models
\textbf{and} forecasting horizons. Furthermore, variations in accuracy
across different forecasting horizons \emph{for each} model is not
observable in Figure 5. The mean error and spread of both models across
different time horizons appears to be constant.
\includegraphics[width=1.05000\textwidth]{Forecast error boxplot.png}\\
\emph{Figure 5: Boxplot illustrating distribution of rolling forecast
errors for different time horizons, h}

@@

Table 2 summarizes the forecast error statistics for both the ARIMA and
Prophet model across the three different time horizons.\\
Across all forecast horizons, the ARIMA model appears to be
overestimating returns. This is evident from the negative mean error(ME)
across \(h = (1,5,20)\). On the contrary, the Prophet model
underestimates the forecasts for one-step and five-step ahead
forecasting horizons. It then overstimates returns when forecasting
twenty-steps ahead. The ME results shown in Table 2 fail to clearly
illustrate which model performs better. Furthermore, the ME values are
relatively small when compared against other error statistics.

@@

The observations made from Table 2 in section 5.2.1 suggest that the
ARIMA and Prophet models have equal predictive accuracy. These
assertions can be formally quantified through a Diebold-Mariano(DM) test
which evaluates whether the models have equal predictive accuracy
against the alternative hypothesis that one model has superior accuracy
over the other. Table 3 reports the results obtained when the rolling
forecast errors were formally compared against each other over different
time horizons.\\
\includegraphics[width=1.05000\textwidth]{dm results.png}\\
\emph{Table 3 Results of the Diebold-Mariano Hypothesis Test}\\
The p-value across all the time horizons for either linear and quadratic
loss functions suggest that we would fail to reject the null hypothesis.
This implies that the predictive accuracy of the models could be equal.
These results formally validate the observations made obtained in
section 5.2.1.

@@

The results of the study show that there is no significant evidence to
conclude that Prophet generates superior forecasts for returns on the
FTSE/JSE Top40 Index when compared against those of an ARIMA model when
the Diebold-Mariano Test is used as a method of evaluation. These
results are consistent for daily, weekly and monthly forecasting
horizons.\\
Nevertheless, the study has flagged the possibility that, on average,
the ARIMA model tends to overestimate forecasted returns on the index
for the forecast horizons under consideration. This is contrary to
Prophet, that, on average, underestimates forecasted returns over daily
and weekly time horizons. This information could be useful for an
analyst engaged in short-term strategies since the choice of the
forecasting model can reflect a particular sentiment or risk appetite.
For example, if the analyst feels like prevailing returns are
underpriced by the market, they might choose to use an ARIMA model,
whereas an opposite sentiment could warrant the use of Prophet.\\
Furthermore, it is worth noting that the merits of Prophet may have not
been fully exposed in this study due to the use of transformed data and
the choice of a forecasting horizon that excluded periods of high
volatility. Nonetheless, it is reassuring that the model has performed
as well as the ARIMA model. This observation could create room for the
application of Prophet in circumstances where flexibility in the model
formulation is desireable.

@@

The challenge of building models that generate superior stock return
forecasts has led to widespread research in the field of quantitative
analysis for decades, but the ARIMA model remains the standard against
which to compare new innovations. This is because of its relative
simplicity and ability to evolve into more complex applications e.g in
the hybrid models considered in the literature review.\\
Prophet has reduced the problem of estimating a time series model to a
curve fittig exercise by defining an additive, parametric function of
time. The functions' parameters are then estimated using Bayesian
estimation methods after which the tool automatically executes forecast
evaluation to optimize its predictive ability.\\
The results of this study suggest that, using a Diebold-Mariano
evaluation, the ARIMA and Prophet forecasting models do not have a
significant difference in predictive accuracy across time horizons of up
to a month. The implications of this are that analysts can then choose
which model to use based on their needs without significantly
compromising forecast accuracy.\\
This study was confined to the analysis of predictive performance during
general market conditions. Further research could explore the predictive
ability of the models during different market conditions. For example,
performance can be compared during bull and bear market conditions or
during periods of high volatility. Additionally, further research can
assess whether prophet would perform better on raw data that has not
been transformed, since the model is built with growth and seasonality
components. This use of which could lead to Prophet models that generate
superior forecasts when compared against alternatives.
